---
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
rm(list=ls())
library(readr)
library(choroplethr)
library(choroplethrMaps)
library(dplyr)
library(ggplot2) 
library(plotly)
library(tmap)
library(sf)
library(tmaptools)
library(gganimate)
library(cowplot)
library(gridExtra)
library(shiny)
library(ggiraph)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# Importing in geojoson file 
my_map <- st_read("~/Documents/mackcross/sci_comm/shapefiles/rural_variables.geojson", quiet=TRUE)
# lab: 
# personal: ~/Documents/mackcross/sci_comm/shapefiles/rural_variables.geojson

# Convert size variable to numeric and remove missing values
my_map$cleaned_rural_density <- as.numeric(my_map$cleaned_rural_density)
my_map$cleaned_rural_mean_sci <- as.numeric(my_map$cleaned_rural_mean_sci)
my_map$cleaned_rural_mean_math <- as.numeric(my_map$cleaned_rural_mean_math)
my_map$cleaned_rural_CountyGraduateRate4yr <- as.numeric(my_map$cleaned_rural_CountyGraduateRate4yr)
my_map$cleaned_rural_CountyGraduateRate5yr <- as.numeric(my_map$cleaned_rural_CountyGraduateRate5yr)
my_map$cleaned_rural_CountySpendingPerStudent <- as.numeric(my_map$cleaned_rural_CountySpendingPerStudent)
my_map$cleaned_rural_Graduate.or.professional.degree.. <- as.numeric(my_map$cleaned_rural_Graduate.or.professional.degree..)
my_map$cleaned_rural_Bachelor.s.degree.. <- as.numeric(my_map$cleaned_rural_Bachelor.s.degree..)
my_map$cleaned_rural_High.school.graduate..includes.equivalency... <- as.numeric(my_map$cleaned_rural_High.school.graduate..includes.equivalency...)

quantile(my_map$cleaned_rural_Graduate.or.professional.degree..)

my_map$HS_Quantile <- cut(my_map$cleaned_rural_High.school.graduate..includes.equivalency..., 
                     breaks = c(17.00, 35.05, 38.00, 41.05, 46.80), 
                     include.lowest = TRUE)

my_map$BA_Quantile <- cut(my_map$cleaned_rural_Bachelor.s.degree.., 
                     breaks = c(7.40, 11.75, 13.50, 14.90, 32.10), 
                     include.lowest = TRUE)

my_map$Grad_Quantile <- cut(my_map$cleaned_rural_Graduate.or.professional.degree.., 
                     breaks = c(2.7, 4.0, 4.8, 5.9, 23.7), 
                     include.lowest = TRUE)


my_map$CountySpending <- cut(my_map$cleaned_rural_CountySpendingPerStudent,
                             breaks = c(12655.18, 15584.22, 16883.50, 18780.58, 33410.00), 
                             include.lowest = TRUE)

my_map$ACT_Sci <- cut(my_map$cleaned_rural_mean_sci,
                             breaks = c(19.01000, 20.13391, 20.64000, 21.16250, 22.63571),
                             include.lowest = TRUE)

my_map$ACT_Math <- cut(my_map$cleaned_rural_mean_math,
                          breaks = c(19.66250, 21.33818, 21.68000, 22.08706, 23.12000),
                          include.lowest = TRUE)

my_map$ForthYear <- cut(my_map$cleaned_rural_CountyGraduateRate4yr,
                          breaks = c(81.9000, 91.1050, 94.0700, 96.0425, 100.0000),
                          include.lowest = TRUE)

my_map$FifthYear <- cut(my_map$cleaned_rural_CountyGraduateRate5yr,
                           breaks = c(88.08000, 93.45750, 95.26600, 97.39333, 100.0000),
                           include.lowest = TRUE)


# Palettes of interest 
# palette = c("#a6bddb", "#3690c0", "#034e7b"), 
# palette = c("#8c510a", "#d8b365", "#7fbf7b")
# palette = c("#edf8fb", "#b3cde3", "#8c96c6", "#88419d")
# c("#f6eff7", "#bdc9e1", "#67a9cf", "#02818a")
# c("#f1eef6", "#bdc9e1", "#74a9cf", "#0570b0")
# c("#edf8fb", "#b2e2e2", "#66c2a4", "#238b45")
# palette = c("#7b3294", "#c2a5cf", "#f7f7f7", "#a6dba0", "#008837")
 palette = c("#edf8fb", "#b3cde3", "#8c96c6", "#8856a7", "#810f7c")
# palette = c("#f6eff7", "#bdc9e1", "#67a9cf", "#1c9099", "#016c59")
# palette = c("#a6611a", "#dfc27d", "#f5f5f5", "#80cdc1", "#018571")
# palette = c("#edf8fb", "#b2e2e2", "#66c2a4", "#2ca25f", "#006d2c")

```


```{r echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#      theme(legend.key.size = unit(2, "cm"), 
#            legend.text = element_text(size = 12), 
#            legend.title = element_text(size = 14, face = "bold")) +


```


# VISUALIZING IOWA EDUCATION

This collection of maps explores some variables related to educational resource allocation and measures used to evaluate student success. 

# School Spending Per Student

```{r echo=FALSE, message=FALSE, warning=FALSE}


ui <- fluidPage(
  navbarPage(
    "Maps",
    tabPanel("Avg. Amount Spent Per Student", 
             plotlyOutput("CountySpending_map"))
  )
)

server <- function(input, output) {
  output$CountySpending_map <- renderPlotly({
    gg <- ggplot() +
      geom_sf(data = my_map, aes(fill = CountySpending,
                                 text = paste("County: ", CountyName, "<br>",
                                             "Avg. Amount Spent: ", 
                                             "$",
                                             round(cleaned_rural_CountySpendingPerStudent)))) +
      scale_fill_manual(values = c("#edf8fb", "#b2e2e2", "#66c2a4", "#238b45"), 
                        name = "Avg. Amount Spent", 
                        na.value = "black") +
      theme(axis.ticks = element_blank(), panel.grid = element_blank(), axis.text = element_blank(), panel.background = element_rect(fill = "white"))
    
    ggplotly(gg, tooltip = "text", 
             hovertemplate = paste(
               "<b>%{text}</b><br><br>",
               "<extra></extra>"
             )
    )
  })
}

shinyApp(ui = ui, server = server)



```


# Levels of Education 
The below two maps provide the levels of education depicted as proportional symbols. They are presented on a choropleth map showing the average amount of money spent per student. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
ui <- fluidPage(
  navbarPage(
    "Maps",
    tabPanel("High School Degrees", 
             plotlyOutput("hs_map")),
    tabPanel("Bachelor's Degrees",
             plotlyOutput("ba_map")), 
    tabPanel("Graduate Degrees", 
             plotlyOutput("grad_map"))
  )
)

server <- function(input, output) {
  output$hs_map <- renderPlotly({
    ggplot() +
      geom_sf(data = my_map, aes(fill = HS_Quantile,
                                 text = paste("County: ", CountyName, "<br>",
                                             "% High School Degrees: ", 
                                             round(cleaned_rural_High.school.graduate..includes.equivalency..., 2), "%"))) +
    scale_fill_manual(values = c("#edf8fb", "#b3cde3", "#8c96c6", "#88419d"), 
                      labels = c("17-34.5 %", "< 37.5%", "< 39.2%", "< 41.4%", "< 46.8%"), 
                      name = "% High School Degrees", 
                      na.value = "black") +
    theme(axis.ticks = element_blank(), panel.grid = element_blank(), axis.text = element_blank(), panel.background = element_rect(fill = "white")) 
})


  output$ba_map <- renderPlotly({
    ggplot()  +
        geom_sf(data = my_map, aes(fill = BA_Quantile,
                                text = paste("County: ", CountyName, "<br>",
                                             "% Bachelor's Degrees: ", 
                                             round(cleaned_rural_Bachelor.s.degree.., 2), "%"))) +
        scale_fill_manual(values = c("#edf8fb", "#b3cde3", "#8c96c6", "#88419d"), 
                          labels = c("7.4-11.2 %", "< 12.9%", "< 14%", "< 15.2%", "< 32.1%"),
                          name = "% Bachelor's Degrees", 
                          na.value = "black") +
        theme(axis.ticks = element_blank(), panel.grid = element_blank(), axis.text = element_blank(), panel.background = element_rect(fill = "white"))  
})

      output$grad_map <- renderPlotly({
    ggplot()  +
        geom_sf(data = my_map, aes(fill = Grad_Quantile,
                                text = paste("County: ", CountyName, "<br>",
                                             "% Graduate Degrees: ", 
                                             round(cleaned_rural_Graduate.or.professional.degree.., 2), "%"))) +
        scale_fill_manual(values = c("#edf8fb", "#b3cde3", "#8c96c6", "#88419d"), 
                          labels = c("2.7 %", "< 4.0%", "< 4.8%", "< 5.9%", "< 23.7%"),
                          name = "% Graduate Degrees", 
                          na.value = "black") +
        theme(axis.ticks = element_blank(), panel.grid = element_blank(), axis.text = element_blank(), panel.background = element_rect(fill = "white")) 
})
  
}

shinyApp(ui = ui, server = server)

```



# Graduation Rate 
The below two maps provide the % of student who graduate from high school in four years for each county depicted as a proportional symbols. They are presented on a choropleth map showing the average amount of money spent per student. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

ui <- fluidPage(
  navbarPage(
    "Maps",
    tabPanel("% 4th Year Graduates", 
             plotlyOutput("ForthYear_map")),
    tabPanel("% 5th Year Graduates",
             plotlyOutput("FifthYear_map")) 

  )
)

server <- function(input, output) {
  output$ForthYear_map <- renderPlotly({
    ggplot() +
      geom_sf(data = my_map, aes(fill = ForthYear,
                                 text = paste("County: ", CountyName, "<br>",
                                             "% 4th Year Graduates: ", 
                                             round(cleaned_rural_CountyGraduateRate4yr, 2), "%"))) +
    scale_fill_manual(values = c("#edf8fb", "#b2e2e2", "#66c2a4", "#238b45"), 
                      labels = c("81.9 - 91%", "< 94%", "< 96%", "<= 100%"), 
                      name = "% 4th Year Graduates", 
                      na.value = "black") +
    theme(axis.ticks = element_blank(), panel.grid = element_blank(), axis.text = element_blank(), panel.background = element_rect(fill = "white")) 
})


  output$FifthYear_map <- renderPlotly({
    ggplot()  +
        geom_sf(data = my_map, aes(fill = FifthYear,
                                text = paste("County: ", CountyName, "<br>",
                                             "% 5th Year Graduates: ", 
                                             round(cleaned_rural_CountyGraduateRate5yr, 2), "%"))) +
        scale_fill_manual(values = c("#edf8fb", "#b2e2e2", "#66c2a4", "#238b45"), 
                          labels = c("88 - 91%", "< 93.46%", "< 95.3%", "<= 100%"),
                          name = "% 5th Year Graduates", 
                          na.value = "black") +
        theme(axis.ticks = element_blank(), panel.grid = element_blank(), axis.text = element_blank(), panel.background = element_rect(fill = "white"))  
})
  
}

shinyApp(ui = ui, server = server)

```



# ACT Math & Science Score Averages 
The below two maps provide the average ACT Math and Science scores for each county depicted as a proportional symbols. They are presented on a choropleth map showing the average amount of money spent per student. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
ui <- fluidPage(
  navbarPage(
    "Maps",
    tabPanel("ACT Science Scores", 
             plotlyOutput("ACT_Sci_map")),
    tabPanel("ACT Math Scores",
             plotlyOutput("ACT_Math_map")) 

  )
)

server <- function(input, output) {
  output$ACT_Sci_map <- renderPlotly({
    ggplot() +
      geom_sf(data = my_map, aes(fill = ACT_Sci,
                                 text = paste("County: ", CountyName, "<br>",
                                             "Avg. ACT Sci Score: ", 
                                             round(cleaned_rural_mean_sci, 2), "%"))) +
    scale_fill_manual(values = c("#f6eff7", "#bdc9e1", "#67a9cf", "#02818a"), 
                      labels = c("17-34.5 %", "< 37.5%", "< 39.2%", "< 41.4%", "< 46.8%"), 
                      name = "Avg. ACT Science Score", 
                      na.value = "black") +
    theme(axis.ticks = element_blank(), panel.grid = element_blank(), axis.text = element_blank(), panel.background = element_rect(fill = "white")) 
})


  output$ACT_Math_map <- renderPlotly({
    ggplot()  +
        geom_sf(data = my_map, aes(fill = ACT_Math,
                                text = paste("County: ", CountyName, "<br>",
                                             "Avg. ACT Math Score: ", 
                                             round(cleaned_rural_mean_math, 2), "%"))) +
        scale_fill_manual(values = c("#f6eff7", "#bdc9e1", "#67a9cf", "#02818a"), 
                          labels = c("7.4-11.2 %", "< 12.9%", "< 14%", "< 15.2%", "< 32.1%"),
                          name = "Avg. ACT Math Score", 
                          na.value = "black") +
        theme(axis.ticks = element_blank(), panel.grid = element_blank(), axis.text = element_blank(), panel.background = element_rect(fill = "white"))  
})
  
}

shinyApp(ui = ui, server = server)
```





