---
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# VISUALIZING IOWA EDUCATION

This collection of maps explores some variables related to educational resource allocation and measures used to evaluate student success. 

```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
rm(list=ls())
library(readr)
library(choroplethr)
library(choroplethrMaps)
library(dplyr)
library(ggplot2) 
library(plotly)
library(tmap)
library(sf)
library(tmaptools)
library(gganimate)
library(cowplot)
library(gridExtra)
library(shiny)
```


```{r echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# Importing in geojoson file 
my_map <- st_read("~/Documents/mackcross/sci_comm/shapefiles/rural_variables.geojson", quiet=TRUE)
# lab: 
# personal: ~/Documents/mackcross/sci_comm/shapefiles/rural_variables.geojson
my_data <- data.frame(my_map)

# Convert size variable to numeric and remove missing values
my_map$cleaned_rural_density <- as.numeric(my_map$cleaned_rural_density)
my_map$cleaned_rural_mean_sci <- as.numeric(my_map$cleaned_rural_mean_sci)
my_map$cleaned_rural_mean_math <- as.numeric(my_map$cleaned_rural_mean_math)
my_map$cleaned_rural_CountyGraduateRate4yr <- as.numeric(my_map$cleaned_rural_CountyGraduateRate4yr)
my_map$cleaned_rural_CountyGraduateRate5yr <- as.numeric(my_map$cleaned_rural_CountyGraduateRate5yr)
my_map$cleaned_rural_CountySpendingPerStudent <- as.numeric(my_map$cleaned_rural_CountySpendingPerStudent)
my_map$cleaned_rural_Graduate.or.professional.degree.. <- as.numeric(my_map$cleaned_rural_Graduate.or.professional.degree..)
my_map$cleaned_rural_Bachelor.s.degree.. <- as.numeric(my_map$cleaned_rural_Bachelor.s.degree..)
my_map$cleaned_rural_High.school.graduate..includes.equivalency... <- as.numeric(my_map$cleaned_rural_High.school.graduate..includes.equivalency...)

# Determine quantiles 
quantile(my_map$cleaned_rural_CountySpendingPerStudent, na.rm = TRUE)

# Set quantiles 
my_map$cleaned_rural_CountySpendingPerStudent <- as.numeric(my_map$cleaned_rural_CountySpendingPerStudent)
my_map$binned_CountySpendingPerStudent <- cut(my_map$cleaned_rural_CountySpendingPerStudent,
                              breaks = c(12655.18, 15584.22, 16883.50, 18780.58, 33410.00),
                              labels = c(1, 2, 3, 4))
my_map$binned_CountySpendingPerStudent <- as.numeric(my_map$binned_CountySpendingPerStudent)


my_map$binned_science <- as.numeric(my_map$cleaned_rural_mean_sci)
my_map$binned_science <- cut(my_map$cleaned_rural_mean_sci,
                             breaks = c(19.01000, 20.13391, 20.64000, 21.16250, 22.63571),
                             labels = c(1, 2, 3, 4))
my_map$binned_science <- as.numeric(my_map$binned_science)


my_map$binned_math <- as.numeric(my_map$cleaned_rural_mean_math)
my_map$binned_math <- cut(my_map$cleaned_rural_mean_math,
                          breaks = c(19.66250, 21.33818, 21.68000, 22.08706, 23.12000),
                          labels = c(1, 2, 3, 4))
my_map$binned_math <- as.numeric(my_map$binned_math)

my_map$binned_grad <- as.numeric(my_map$cleaned_rural_CountyGraduateRate4yr)
my_map$binned_grad <- cut(my_map$cleaned_rural_CountyGraduateRate4yr,
                          breaks = c(81.9000, 91.1050, 94.0700, 96.0425, 100.0000),
                          labels = c(1, 2, 3, 4))
my_map$binned_grad <- as.numeric(my_map$binned_grad)


my_map$binned_grad5 <- as.numeric(my_map$cleaned_rural_CountyGraduateRate5yr)
my_map$binned_grad5 <- cut(my_map$cleaned_rural_CountyGraduateRate5yr,
                           breaks = c(88.08000, 93.45750, 95.26600, 97.39333, 100.0000),
                           labels = c(1, 2, 3, 4))
my_map$binned_grad5 <- as.numeric(my_map$binned_grad5)


my_map$binned_phd <- as.numeric(my_map$cleaned_rural_Graduate.or.professional.degree..)
my_map$binned_phd <- cut(my_map$cleaned_rural_Graduate.or.professional.degree..,
                         breaks = c(2.7, 4.0, 4.8, 5.9, 23.7),
                         labels = c(1, 2, 3, 4))
my_map$binned_phd <- as.numeric(my_map$binned_phd)


my_map$binned_ba <- as.numeric(my_map$cleaned_rural_Bachelor.s.degree..)
my_map$binned_ba <- cut(my_map$cleaned_rural_Bachelor.s.degree..,
                        breaks = c(7.40, 11.75, 13.50, 14.90, 32.10),
                        labels = c(1, 2, 3, 4))
my_map$binned_ba <- as.numeric(my_map$binned_ba)

# Palettes of interest 
# palette = c("#a6bddb", "#3690c0", "#034e7b"), 
# palette = c("#8c510a", "#d8b365", "#7fbf7b"),
# palette = c("#7b3294", "#c2a5cf", "#f7f7f7", "#a6dba0", "#008837")
# palette = c("#edf8fb", "#b3cde3", "#8c96c6", "#8856a7", "#810f7c")
# palette = c("#f6eff7", "#bdc9e1", "#67a9cf", "#1c9099", "#016c59")
# palette = c("#a6611a", "#dfc27d", "#f5f5f5", "#80cdc1", "#018571")
# palette = c("#edf8fb", "#b2e2e2", "#66c2a4", "#2ca25f", "#006d2c")
```


# ACT Math & Science Score Averages 
The below two maps provide the average ACT Math and Science scores for each county depicted as a proportional symbols. They are presented on a choropleth map showing the average amount of money spent per student. 

```{r, echo=FALSE}
ui <- fluidPage(
  navbarPage(
    "Maps",
    tabPanel("Science Scores", 
             tmapOutput("map_science")),
    tabPanel("Math Scores",
             tmapOutput("map_math"))
  )
)

server <- function(input, output) {
  output$map_science <- renderTmap({
    tm_shape(my_map) +
      tm_fill("cleaned_rural_mean_sci", 
              style = "quantile", 
              n = 5, 
              palette = c("#f6eff7", "#bdc9e1", "#67a9cf", "#1c9099", "#016c59"), 
              title = "Avg. ACT Science Score") +
      tm_symbols(size = "binned_CountySpendingPerStudent",
                 col = "black", 
                 alpha = .8, 
                 scale = 1.75)
  })
  
  output$map_math <- renderTmap({
    tm_shape(my_map) +
      tm_fill("cleaned_rural_mean_math", 
              style = "quantile", 
              n = 5, 
              palette = c("#f6eff7", "#bdc9e1", "#67a9cf", "#1c9099", "#016c59"), 
              title = "Avg. ACT Math Score") +
      tm_symbols(size = "binned_CountySpendingPerStudent",
                 col = "black",
                 alpha = .8, 
                 scale = 1.75)
  })
}
shinyApp(ui = ui, server = server)
```



# Graduation Rate 
The below two maps provide the % of student who graduate from high school in four years for each county depicted as a proportional symbols. They are presented on a choropleth map showing the average amount of money spent per student. 

```{r, echo=FALSE}
ui <- fluidPage(
  navbarPage(
    "Maps",
    tabPanel("% 4yr Grad", 
             tmapOutput("map_grad")),
    tabPanel("% 5yr Grad",
             tmapOutput("map_grad5"))
  )
)

server <- function(input, output) {
  output$map_grad <- renderTmap({
    tm_shape(my_map) +
      tm_fill("cleaned_rural_CountyGraduateRate4yr", 
              style = "quantile", 
              n = 5, 
              palette = c("#edf8fb", "#b3cde3", "#8c96c6", "#8856a7", "#810f7c"), 
              title = "4th Year Graduate Rate") +
      tm_bubbles(size = "binned_CountySpendingPerStudent",
                 col = "black", 
                 alpha = .8, 
                 scale = 1.75) +
      tm_legend(outside = TRUE, 
                outside.position = "left", 
                bg.color = "transparent", 
                frame = FALSE)
  })
  
  output$map_grad5 <- renderTmap({
    tm_shape(my_map) +
      tm_fill("cleaned_rural_CountyGraduateRate5yr", 
              style = "quantile", 
              n = 5, 
              palette = c("#edf8fb", "#b3cde3", "#8c96c6", "#8856a7", "#810f7c"), 
              title = "5th Year Graduate Rate") +
      tm_bubbles(size = "binned_CountySpendingPerStudent",
                 col = "black",
                 alpha = .8, 
                 scale = 1.75) +
      tm_legend(outside = TRUE, 
                outside.position = "left", 
                bg.color = "transparent", 
                frame = FALSE)
  })
}

shinyApp(ui = ui, server = server)
```



# Levels of Education 
The below two maps provide the levels of education depicted as proportional symbols. They are presented on a choropleth map showing the average amount of money spent per student. 

```{r, echo=FALSE}
ui <- fluidPage(
  navbarPage(
    "Maps",
    tabPanel("% High School Degree",
             tmapOutput("binned_hs")), 
    tabPanel("% Bachelors Degree", 
             tmapOutput("binned_ba")),
    tabPanel("% Graduate Degree",
             tmapOutput("binned_phd"))
  )
)

server <- function(input, output) {
  output$binned_phd <- renderTmap({
  
tm_shape(my_map) +
  tm_fill("cleaned_rural_Graduate.or.professional.degree..", style = "quantile", n = 5, palette = c("#edf8fb", "#b2e2e2", "#66c2a4", "#2ca25f", "#006d2c"), title = "% Graduate Degree Holders") +
  tm_bubbles(size = "binned_CountySpendingPerStudent",
             col = "black",
             n = 3, 
             alpha = .8, 
             scale = 1.75) +
  tm_legend(outside = TRUE, 
            outside.position = "left", 
            bg.color = "transparent", 
            frame = FALSE)

  })
  
  output$binned_ba <- renderTmap({

tm_shape(my_map) +
  tm_fill("cleaned_rural_Bachelor.s.degree..", style = "quantile", n = 5, palette = c("#edf8fb", "#b2e2e2", "#66c2a4", "#2ca25f", "#006d2c"), title = "% Bachelor's Degree Holders") +
  tm_bubbles(size = "binned_CountySpendingPerStudent",
             col = "black",
             alpha = .8,
             scale = 1.75) +
  tm_legend(outside = TRUE, 
            outside.position = "left", 
            bg.color = "transparent", 
            frame = FALSE)

  })
  
    output$binned_hs <- renderTmap({

tm_shape(my_map) +
  tm_fill("cleaned_rural_High.school.graduate..includes.equivalency...", style = "quantile", n = 5, palette = c("#edf8fb", "#b2e2e2", "#66c2a4", "#2ca25f", "#006d2c"), title = "% High School Degree Holders") +
  tm_bubbles(size = "binned_CountySpendingPerStudent",
             col = "black",
             alpha = .8,
             scale = 1.75) +
  tm_legend(outside = TRUE, 
            outside.position = "left", 
            bg.color = "transparent", 
            frame = FALSE)

  })
}

shinyApp(ui = ui, server = server)
```

# END

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# Determine quantiles 
# quantile(my_map$cleaned_rural_mean_sci, na.rm = TRUE)

# Set quantiles 
# my_map$binned_science <- as.numeric(my_map$cleaned_rural_mean_sci)
# my_map$binned_science <- cut(my_map$cleaned_rural_mean_sci,
                             # breaks = c(19.01000, 20.13391, 20.64000, 21.16250, 22.63571),
                             # labels = c(1, 2, 3, 4))
# my_map$binned_science <- as.numeric(my_map$binned_science)


# tm_shape(my_map) +
#   tm_fill("cleaned_rural_CountySpendingPerStudent", style = "quantile", n = 5, palette = "Greys", title = "Avg. Spending per Student") +
#  tm_bubbles(size = "binned_science",
#             col = "#034e7b", 
#             alpha = .8, 
#             title.size ="Avg. ACT Science Score", 
#             legend.size.is.portrait=TRUE,
#             sizes.legend.labels=c("<= 20.1","<= 20.6","<= 21.2","<= 22.6"), 
#             scale = 1.75) +
  #tm_layout(frame = FALSE, bg.color = "black") +
#  tm_legend(outside = TRUE, 
#            outside.position = "left", 
#            bg.color = "transparent", 
#            frame = FALSE)

###############################################################################

# Determine quantiles 
# quantile(my_map$cleaned_rural_mean_math, na.rm = TRUE)

# Set quantiles 
#my_map$binned_math <- as.numeric(my_map$cleaned_rural_mean_math)
#my_map$binned_math <- cut(my_map$cleaned_rural_mean_math,
#                             breaks = c(19.66250, 21.33818, 21.68000, 22.08706, 23.12000),
#                             labels = c(1, 2, 3, 4))
#my_map$binned_math <- as.numeric(my_map$binned_math)


#tm_shape(my_map) +
#  tm_fill("cleaned_rural_CountySpendingPerStudent", style = "quantile", n = 5, palette = "Greys", title = "Avg. Spending per Student") +
#  tm_bubbles(size = "binned_math",
#             col = "#034e7b",
#             alpha = .8, 
#             title.size ="Avg. ACT Math Score", 
#             legend.size.is.portrait=TRUE,
#             sizes.legend.labels=c("<= 21.3","<= 21.7","<= 22.1","<= 23.1"), 
#             scale = 1.75) +
#  tm_layout(frame = FALSE, bg.color = "grey") +
#  tm_legend(outside = TRUE, 
#            outside.position = "left", 
#            bg.color = "transparent", 
#            frame = FALSE)


```



```{r, echo=FALSE}
# Determine quantiles 
# quantile(my_map$cleaned_rural_CountyGraduateRate4yr, na.rm = TRUE)

# Set quantiles 
#my_map$binned_grad <- as.numeric(my_map$cleaned_rural_CountyGraduateRate4yr)
#my_map$binned_grad <- cut(my_map$cleaned_rural_CountyGraduateRate4yr,
#                             breaks = c(81.9000, 91.1050, 94.0700, 96.0425, 100.0000),
#                             labels = c(1, 2, 3, 4))
#my_map$binned_grad <- as.numeric(my_map$binned_grad)


#tm_shape(my_map) +
#  tm_fill("cleaned_rural_CountySpendingPerStudent", style = "quantile", n = 5, palette = "Greys", title = "Avg. Spending per Student") +
#  tm_bubbles(size = "binned_grad",
#             col = "#6a51a3", 
#             n = 3, 
#             alpha = .8, 
#             title.size ="4yr Graduation Rate", 
#             legend.size.is.portrait=TRUE,
#             sizes.legend.labels=c("<= 91%","<= 94%","<= 96%","<= 100%"), 
#             scale = 1.75) +
  #tm_layout(frame = FALSE, bg.color = "black") +
#  tm_legend(outside = TRUE, 
#            outside.position = "right", 
#            bg.color = "transparent", 
#            frame = FALSE)

# Determine quantiles 
# quantile(my_map$cleaned_rural_CountyGraduateRate5yr, na.rm = TRUE)

# Set quantiles 
#my_map$binned_grad5 <- as.numeric(my_map$cleaned_rural_CountyGraduateRate5yr)
#my_map$binned_grad5 <- cut(my_map$cleaned_rural_CountyGraduateRate5yr,
#                             breaks = c(88.08000, 93.45750, 95.26600, 97.39333, 100.0000),
#                             labels = c(1, 2, 3, 4))
#my_map$binned_grad5 <- as.numeric(my_map$binned_grad5)


#tm_shape(my_map) +
#  tm_fill("cleaned_rural_CountySpendingPerStudent", style = "quantile", n = 5, palette = "Greys", title = "Avg. Spending per Student") +
#  tm_bubbles(size = "binned_grad5",
#             col = "#6a51a3", 
#             n = 3, 
#             alpha = .8, 
#             title.size ="5yr Graduation Rate", 
#             legend.size.is.portrait=TRUE,
#             sizes.legend.labels=c("<= 93.5%","<= 95%","<= 97.5%","<= 100%"), 
#             scale = 1.75) +
  #tm_layout(frame = FALSE, bg.color = "black") +
#  tm_legend(outside = TRUE, 
#            outside.position = "right", 
#            bg.color = "transparent", 
#            frame = FALSE)

```



